---
title: "Mid-Atlantic Regression --- Proposition 4 submodels (reference copy)"
keep-md: true
format:
  pdf:
    toc: true
    number-sections: true
    keep-tex: true
    header-includes: |
      \usepackage{hyperref}
      \hypersetup{
        pdfpagemode=UseOutlines, % open bookmarks panel
        bookmarksopen=true,      % unfold top-level bookmarks
        bookmarksopenlevel=10,   % unfold all levels
        bookmarksnumbered=true   % number bookmarks
      }
---

---

__Code version: `r packageVersion("sphm")`__

> **Reference copy.** This is Shogo Kato's original document for the four Proposition 4
> submodels, kept only so its output can be compared against the merged
> `reproduce_midatlantic.qmd`. It calls the preserved `obsolete_` implementation rather
> than the package API. Do not develop against it.

```{r prep}
#| include: false
#| cache: false
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(R.matlab)
library(ggplot2)
library(sphm)
library("R.matlab")
library(tibble)
library(dplyr)
library(patchwork)
library(kableExtra)

# The Proposition 4 code this document was written against is preserved verbatim in the
# package as R/obsolete_Mobius_prop4.R, under obsolete_-prefixed names. See that file's
# header, and tests/testthat/test-prop4_vs_obsolete.R.
mobius_SvMF            <- sphm:::obsolete_mobius_SvMF
mobius_link            <- sphm:::obsolete_mobius_link
as_mobius_link_prop4ii <- sphm:::obsolete_as_mobius_link_prop4ii
```

```{r Residual Plotting Function}
#| include: false
viridis_switch <- function(continuous) {
  if (continuous) {
    return(scale_color_viridis_c)
  } else {
    return(scale_color_viridis_d)
  }
}

soloplt <- function(rresids, covars, extraaes = list(NULL), continuous = TRUE, basemapping = aes(x = r1, y = r2)){
  colorfun <- viridis_switch(continuous = continuous)
  bind_cols(rresids, covars) %>%
    ggplot(modifyList(basemapping, extraaes)) +
    geom_point() +
    colorfun() +
    # scale_y_continuous(transform = scales::transform_pseudo_log(sigma = pi/32, base = exp(1))) +
    # scale_x_continuous(transform = scales::transform_pseudo_log(sigma = pi/32, base = exp(1))) +
    geom_vline(xintercept = 0, lty = "dotted", col = "blue") +
    geom_hline(yintercept = 0, lty = "dotted", col = "blue") +
    coord_equal()
}
# soloplt(small_LinEuc$rresid, sphdf, aes(col = Depth), continuous = TRUE)
```

```{r Tools for ggally with circle}
#| include: false
# Create data for the unit circle
theta <- seq(0, 2 * pi, length.out = 100)
circle_df <- data.frame(x = cos(theta), y = sin(theta))

custom_ggally_points <- function(data, mapping, ...) {
  ggplot(data, mapping) +
    geom_point(...) +
    geom_path(data = circle_df, aes(x = x, y = y), inherit.aes = FALSE, color = "grey") +
    coord_fixed()
}
```

# Data

```{r data}
#| include: false
raw <- readMat("midatlantic.mat")
rownames(raw$X) <- paste0("Y", 1:3)
rownames(raw$Y) <- paste0("X", 1:3) #X3 runs from north pole towards south pole
fulldf <- bind_cols(t(raw$X), t(raw$Y)) %>%
  mutate(westedge = Y2 < -0.3)
```

# Angular Distance Between Observations
```{r}
Y <- fulldf %>%
  select(starts_with("Y")) %>%
  as.matrix()
ang_dist <- acos(pmin(Y %*% t(Y), 1)) # pmin clamps dot products to <=1, preventing NaN from acos due to floating-point rounding
tibble::enframe(ang_dist[lower.tri(ang_dist)], value = "ang_dist") %>%
  ggplot() +
  geom_histogram(aes(x = ang_dist), bins = 30) +
  geom_rug(aes(x = ang_dist)) +
  scale_x_continuous(name = "Angular Distance (radians)")
```

# Euc Covariates
```{r}
xe <- fulldf %>% 
  select(westedge)
xestd <- xe %>% scale() %>% as_tibble()
```

# Our SvMF Models
```{r ours}
#| include: false
#| cache: false
# >>> PROPOSITION 4(i)/(ii) MODELS: MODIFIED
y_data <- fulldf %>% select(starts_with("Y")) %>% as.matrix()
xs_data <- fulldf %>% select(starts_with("X")) %>% as.matrix()
xe_all <- xe %>% as.matrix()

# Proposition 4(i): spherical covariate only, Bs = beta_s I_(p-1).
# This is the strict Möbius-link model in Proposition 4(i), so qe = 0.
mod_prop4i <- mobius_SvMF(
  y = y_data,
  xs = xs_data,
  xe = NULL,
  type = "Prop4i",
  det_constraint = "orthogonal",
  G01behaviour = "free"
)

# Proposition 4(i)-constrained spherical component plus all Euclidean
# covariates.  Proposition 4(i) itself is stated for qe=0; this extended
# model keeps Bs = beta_s I_(p-1) for the spherical part and adds the
# same LinEuc term used in the other Euclidean model.
mod_prop4i_all <- mobius_SvMF(
  y = y_data,
  xs = xs_data,
  xe = xe_all,
  mean = mod_prop4i$mean,
  type = "Prop4i",
  intercept = TRUE,
  det_constraint = "orthogonal",
  G01behaviour = "free"
)

# Proposition 4(ii): spherical covariate only, Bs = I_(p-1).
mod_sph <- mobius_SvMF(
  y = y_data,
  xs = xs_data,
  xe = NULL,
  type = "Prop4ii",
  det_constraint = "orthogonal",
  G01behaviour = "free"
)

# Proposition 4(ii) spherical part plus all Euclidean covariates in this analysis.
mod_all <- mobius_SvMF(
  y = y_data,
  xs = xs_data,
  xe = xe_all,
  type = "Prop4ii",
  intercept = TRUE,
  det_constraint = "orthogonal",
  G01behaviour = "free"
)

# Retain the original object name for the existing downstream analysis.
mod <- mod_all
# <<< PROPOSITION 4(i)/(ii) MODELS
```

```{r Proposition 4 model comparison}
# >>> PROPOSITION 4(i)/(ii) MODEL COMPARISON: ADDED
bind_rows(
  `Proposition 4(i): spherical only` = tibble(
    k = mod_prop4i$k, DoF = mod_prop4i$DoF,
    logLik = mod_prop4i$lLik, AIC = mod_prop4i$AIC
  ),
  `Proposition 4(i): spherical + all Euclidean` = tibble(
    k = mod_prop4i_all$k, DoF = mod_prop4i_all$DoF,
    logLik = mod_prop4i_all$lLik, AIC = mod_prop4i_all$AIC
  ),
  `Proposition 4(ii): spherical only` = tibble(
    k = mod_sph$k, DoF = mod_sph$DoF,
    logLik = mod_sph$lLik, AIC = mod_sph$AIC
  ),
  `Proposition 4(ii): spherical + all Euclidean` = tibble(
    k = mod_all$k, DoF = mod_all$DoF,
    logLik = mod_all$lLik, AIC = mod_all$AIC
  ),
  .id = "Model"
)

# Proposition 4(i) isotropic spherical scale.  beta_s = 1 corresponds to
# Proposition 4(ii); phi = (1-beta_s)/(1+beta_s).
bind_rows(
  `Proposition 4(i): spherical only` = tibble(
    beta_s = mod_prop4i$beta_s,
    phi = mod_prop4i$phi,
    psi_norm = sqrt(sum(mod_prop4i$psi^2))
  ),
  `Proposition 4(i): spherical + all Euclidean` = tibble(
    beta_s = mod_prop4i_all$beta_s,
    phi = mod_prop4i_all$phi,
    psi_norm = sqrt(sum(mod_prop4i_all$psi^2))
  ),
  .id = "Model"
)

bind_rows(
  `Proposition 4(i): spherical only` = as_tibble_row(
    setNames(as.list(mod_prop4i$a), paste0("a", seq_along(mod_prop4i$a)))
  ),
  `Proposition 4(i): spherical + all Euclidean` = as_tibble_row(
    setNames(as.list(mod_prop4i_all$a), paste0("a", seq_along(mod_prop4i_all$a)))
  ),
  `Proposition 4(ii): spherical only` = as_tibble_row(
    setNames(as.list(mod_sph$a), paste0("a", seq_along(mod_sph$a)))
  ),
  `Proposition 4(ii): spherical + all Euclidean` = as_tibble_row(
    setNames(as.list(mod_all$a), paste0("a", seq_along(mod_all$a)))
  ),
  .id = "Model"
)
# <<< PROPOSITION 4(i)/(ii) MODEL COMPARISON
```

# Pretty Data Plot
```{r}
ymean <- colMeans(fulldf %>% select(starts_with("Y")))
ymean <- ymean/sqrt(sum(ymean^2))
#project to equator
ymeanproj <- c(ymean[1:2], 0)
ymeanproj/vnorm(ymeanproj)
target <- c(1,0,0)
prettyrot <- parallel_transport_mat(ymeanproj, target)
prettyY <- (fulldf %>%
  select(Y1, Y2, Y3) %>%
  as.matrix()) %*% t(prettyrot)
colnames(prettyY) <- paste0("Y", 1:3)
prettyX <- (fulldf %>%
  select(X1, X2, X3) %>%
  as.matrix()) %*% t(prettyrot)
colnames(prettyX) <- paste0("X", 1:3)
prettyXmean <- colMeans(prettyX)
prettyXmean <- prettyXmean/vnorm(prettyXmean)
# denomloc <- drop(prettyrot %*% as_mobius_link_cann(mod$mean)$Qs[,1])
plotdata <- bind_cols(prettyY, prettyX, westedge = fulldf$westedge) %>%
  as_tibble() %>%
  ggplot() +
  geom_segment(aes(x=X2, y=X3, xend=Y2, yend=Y3), lty = "dashed", col = "grey") +
  geom_point(aes(x=Y2, y=Y3), shape = 4) +
  geom_point(aes(x=X2, y=X3, fill = westedge), show.legend = FALSE, shape = 21) +
  geom_path(data = circle_df, aes(x = x, y = y), inherit.aes = FALSE, color = "grey") +
  scale_fill_manual(values = c("black", "white")) +
  theme_void() +
  coord_fixed() 
  # mark mean of spherical covariate (obtained earlier)
  # annotate("point", x = prettyXmean[2], y = prettyXmean[3], shape = 24, size = 2)
  # annotate("point", x = denomloc[2], y = denomloc[3], shape = 3, size = 4, col = "blue")
plotdata
```

```{r Paine Plot Resids and Dists}
#| include: false
calc_Paine_resids <- function(mod){
  mod$rresids <- rotated_resid(t(mod$Z), t(mod$predmean), base = north_pole(3))[,-1]
  mod$dists <- acos(rowSums(t(mod$Z) * t(mod$predmean)))
  colnames(mod$rresids) <- c("r1", "r2")
  return(mod)
}

predplot_ESAG <- function(esagmod){
  nthproj <- (diag(3) - north_pole(3) %*% t(north_pole(3)))
  obspred <- bind_cols(obs = as_tibble(t(nthproj %*% prettyrot %*% esagmod$Qmat %*% esagmod$Z)[,-1]),
                       pred = as_tibble(t(nthproj %*% prettyrot %*% esagmod$Qmat %*% esagmod$predmean)[,-1]) %>% rename_with(.fn = ~ paste0("p", .x)),
                       .name_repair = "unique")
  obspred %>%
    as_tibble() %>%
    ggplot() +
    geom_segment(aes(x = pV1, y = pV2, xend = V1, yend = V2),
                 arrow = grid::arrow(length = unit(0.01, "npc"))) +
    geom_path(data = circle_df, aes(x = x, y = y), inherit.aes = FALSE, color = "grey") +
    coord_fixed() +
    theme_void()
}
```

# Predictions
Four submodels are fitted above: the strict Proposition 4(i) spherical-only
Möbius-link model, the Proposition 4(i)-constrained spherical model with all
Euclidean covariates, the Proposition 4(ii) spherical-only model, and the
Proposition 4(ii) spherical model with all Euclidean covariates.  The
Proposition 4(i) result itself is stated for no Euclidean covariates; the
second model is an extension that keeps the restriction
$B_s=\beta_s I_{p-1}$ on the spherical component while adding the same
linear Euclidean term used in the Proposition 4(ii)+Euclidean model.  The
object `mod` below continues to refer to the Proposition 4(ii) model
containing all Euclidean covariates, preserving the structure of the
original downstream analysis.

```{r}
# >>> PROPOSITION 4(ii) + OPTIONAL EUCLIDEAN COVARIATES: ADDED
# Construct a representative canonical parameterisation for the existing
# tables and plotting code.  Here Bs is fixed to the identity.
prop4ii_as_cann <- function(mean) {
  mean <- as_mobius_link_prop4ii(mean)
  p <- nrow(mean$Rtilde0)

  if (is.null(mean$P)) {
    return(mobius_link_cann(
      P = mean$Rtilde0,
      Bs = diag(p - 1),
      Qs = diag(p),
      check = FALSE
    ))
  }

  Qs <- t(mean$Rtilde0) %*% mean$P
  Qe <- matrix(0, nrow(mean$Qe_star) + 1, p)
  Qe[1, 1] <- 1
  Qe[-1, -1] <- mean$Qe_star

  mobius_link_cann(
    P = mean$P,
    Bs = diag(p - 1),
    Qs = Qs,
    Be = diag(mean$Be),
    Qe = Qe,
    ce = 1,
    check = FALSE
  )
}

cann <- prop4ii_as_cann(mod$mean)
cann
# <<< PROPOSITION 4(ii) + OPTIONAL EUCLIDEAN COVARIATES
```

Lets try to interpret the fitted link.

```{r}
cann$Bs %*% t(cann$Qs[,-1])
cann$Be %*% t(cann$Qe[,-1])
```

For the spherical component, $B_s=I_2$ is fixed by Proposition 4(ii), so
the first matrix describes the representative orientation $R_s$ rather than
axis-specific spherical scaling.  The second matrix describes the estimated
Euclidean contribution.

```{r}
cann$Qs[,1]
```

The displayed first column of $R_s$ is based on the representative
decomposition $\widetilde R_0=B_0R_s^\top$ used for interpretation.

## Table of Estimates
```{r Estimates Table, results='asis'}
df <- cbind("diag($B_s$)" = diag(cann$Bs),
      "diag($B_e$)" = diag(cann$Be),
      tQe = t(cann$Qe[-1,-1])) %>%
  as.data.frame()
row.names(df) <- c("$t_2$", "$t_3$")
mykbl <- kbl(df, booktabs = TRUE, position = "!h", escape = FALSE, format = "latex", digits = 2) %>%
  add_header_above(c(" "=3, "$R_e^{\\\\top}$" = 2), escape = FALSE)
mykbl
```

## Plot

```{r Projected Predictions Plot}
#| include: false
predplot <- function(svmfresult){
  ystd <- standardise_sph(svmfresult$y, rotation = prettyrot)
  projmat <- (diag(3) - north_pole(3) %*% t(north_pole(3)))
  obs <- ystd %*% projmat[,-1]
  pred <- svmfresult$pred %*% t(attr(ystd, "std_rotation")) %*% projmat[,-1]

  obspred <- bind_cols(obs = as.data.frame(obs),
            pred = as.data.frame(pred) %>% rename_with(.fn = ~ paste0("p", .x)),
            cov = as.data.frame(prettyX) %>% rename_with(.fn = ~ paste0("p", .x)),
            .name_repair = "unique")

  outplot <- obspred %>%
    as_tibble() %>%
    cbind(fulldf) %>%
    mutate(angle = acos(svmfresult$pred %*% svmfresult$G0[,1])) %>%
    ggplot() +
    # geom_point(aes(x=pX2, y=pX3), shape = 3) +
    # geom_segment(aes(x=pX2, y=pX3, xend=pV1, yend=pV2), data = , lty = "dashed", col = "grey") +
    geom_segment(aes(x = pV1, y = pV2, xend = V1, yend = V2),
                 arrow = grid::arrow(length = unit(0.01, "npc"))) +
    geom_path(data = circle_df, aes(x = x, y = y), inherit.aes = FALSE, color = "grey") +
    coord_fixed() +
    theme_void()

  # mark predictions for mean xs
  # >>> PROPOSITION 4(ii) + OPTIONAL EUCLIDEAN COVARIATES: MODIFIED
  nth <- attr(ystd, "std_rotation") %*% c(0,0,1)
  xe_nat <- matrix(c(0, 1), ncol = 1)
  colnames(xe_nat) <- colnames(svmfresult$xe)
  xs <- colMeans(svmfresult$xs)
  xs <- xs/vnorm(xs)
  eastwestedge <- mobius_link(
    xs = rbind(xs, xs),
    xe = xe_nat,
    param = svmfresult$mean
  ) %*% t(attr(ystd, "std_rotation"))
  # <<< PROPOSITION 4(ii) + OPTIONAL EUCLIDEAN COVARIATES
  outplot <- outplot +
    annotate("point", x = eastwestedge[1, 2], y = eastwestedge[1, 3], fill = "black", shape = 24, size = 2) +
    annotate("point", x = eastwestedge[2, 2], y = eastwestedge[2, 3], fill = "white", shape = 24, size = 2)
  
  # mark prediction when xs = rs1
  # rs1 <- prop4ii_as_cann(svmfresult$mean)$Qs[,1]
  # eastwestedge_rs1 <- mobius_link(xs = rbind(rs1, rs1),
  #        xe = tmpxe,
  #        param = svmfresult$mean) %*% t(attr(ystd, "std_rotation"))
  # outplot <- outplot +
  #   annotate("point", x = eastwestedge_rs1[1, 2], y = eastwestedge_rs1[1, 3], fill = "black", shape = 24, size = 4) +
  #   annotate("point", x = eastwestedge_rs1[2, 2], y = eastwestedge_rs1[2, 3], fill = "white", shape = 24, size = 4)
  
  # mark columns of Qs on the plot
  prettyQs <- attr(ystd, "std_rotation") %*% prop4ii_as_cann(svmfresult$mean)$Qs
  outplot <- outplot +
    annotate("point", x = prettyQs[2,1], y = prettyQs[3,1], shape = 3, size = 2, col = switch(1+(prettyQs[1,1] > 0), "grey", "black")) +
    annotate("point", x = prettyQs[2,2], y = prettyQs[3,2], shape = 4, size = 2, col = switch(1+(prettyQs[1,2] > 0), "grey", "black")) +
    annotate("point", x = prettyQs[2,3], y = prettyQs[3,3], shape = 8, size = 2, col = switch(1+(prettyQs[1,3] > 0), "grey", "black"))
  
if (!is.null(svmfresult$G0)){
  G0proj <- (diag(3) - north_pole(3) %*% t(north_pole(3))) %*% attr(ystd, "std_rotation") %*% svmfresult$G0
  mnaxes_proj <- (diag(3) - north_pole(3) %*% t(north_pole(3))) %*% attr(ystd, "std_rotation") %*% prop4ii_as_cann(svmfresult$mean)$P
  eps = 0.2
  outplot <- outplot + 
  annotate("segment", x = G0proj[2,1], y = G0proj[3,1],
           xend = G0proj[2,1] + eps*G0proj[2,2], 
           yend = G0proj[3,1] + eps*G0proj[3,2], col = "grey",
           lty = "solid") +
  annotate("segment", x = G0proj[2,1], y = G0proj[3,1],
           xend = G0proj[2,1] + eps*G0proj[2,3], 
           yend = G0proj[3,1] + eps*G0proj[3,3], col = "grey",
           lty = "31") +
  
  annotate("segment", x = mnaxes_proj[2,1], y = mnaxes_proj[3,1],
           xend = mnaxes_proj[2,1] + eps*mnaxes_proj[2,2], 
           yend = mnaxes_proj[3,1] + eps*mnaxes_proj[3,2], col = "black",
           lty = "solid") +
  annotate("segment", x = mnaxes_proj[2,1], y = mnaxes_proj[3,1],
           xend = mnaxes_proj[2,1] + eps*mnaxes_proj[2,3], 
           yend = mnaxes_proj[3,1] + eps*mnaxes_proj[3,3], col = "black",
           lty = "31") 
}
return(outplot)
}
plotours <- predplot(mod)
plotours
```


